---
title: "F23_Project_Final"
author: "Ethan Carter"
date: "2023-11-30"
output: html_document
---

```{r setup, include=FALSE}
raw_data <- read.csv("C:/Users/ethan/OneDrive/Desktop/New folder/Semester 15/STAT 512/Data/spotify-2023.csv", header =TRUE, sep=",")
raw_data
```




KEY FEATURES:

- track_name: Name of the song 
- artist(s)_name: Name of the artist(s) of the song 
- artist_count: Number of artists contributing to the song 

- released_year: Year when the song was released 
- released_month: Month when the song was released 
- released_day: Day of the month when the song was released

- in_spotify_playlists: Number of Spotify playlists the song is included in 
- in_spotify_charts: Presence and rank of the song on Spotify charts

- streams: Total number of streams on Spotify

- in_apple_playlists: Number of Apple Music playlists the song is included in 
- in_apple_charts: Presence and rank of the song on Apple Music charts

- in_deezer_playlists: Number of Deezer playlists the song is included in 
- in_deezer_charts: Presence and rank of the song on Deezer charts

- in_shazam_charts: Presence and rank of the song on Shazam charts

- bpm: Beats per minute, a measure of song tempo 
- key: Key of the song 
- mode: Mode of the song (major or minor) 
- danceability_%: Percentage indicating how suitable the song is for dancing 
- valence_%: Positivity of the song's musical content 
- energy_%: Perceived energy level of the song 
- acousticness_%: Amount of acoustic sound in the song 
- instrumentalness_%: Amount of instrumental content in the song 
- liveness_%: Presence of live performance elements 
- speechiness_%: Amount of spoken words in the song 

```{r cars}
#install.packages("base")
#install.packages("dplyr")
library("base")
library("dplyr")

raw_dataframe <- data.frame(raw_data)
raw_dataframe

## Uncomment for the raw variables
track_name <- raw_dataframe$track_name

spotify_charts <- raw_dataframe$in_spotify_charts
apple_charts <- raw_dataframe$in_apple_charts

bpm <- raw_dataframe$bpm
danceability <- raw_dataframe$danceability_.
valence <- raw_dataframe$valence_.
energy <- raw_dataframe$energy_.
streams <- raw_dataframe$streams
playlists <- raw_dataframe$in_spotify_playlists


### Mode would need to be translated to a categorical variable
#https://www.statology.org/r-append-to-vector-in-loop/
mode <- c()
#count <- 0
for (mode_. in raw_dataframe$mode){
  if (mode_. == 'Major'){
    mode <- c(mode,1)
    #print (1)
  }
  else if (mode_. == 'Minor'){
    mode <- c(mode,0)
    #print(0)
  }
  else {
    mode <- c(mode," ")
  }
  #count <- count + 1
}
#mode 
#count

unprocessed_dataframe <- data.frame(bpm,danceability,energy,mode,spotify_charts)
processed_dataframe <- na.omit(unprocessed_dataframe)
#processed_dataframe <- processed_dataframe[apply(processed_dataframe!=0,1,all),]
processed_dataframe <- processed_dataframe %>% filter(processed_dataframe$spotify_charts != 0)

## Uncomment for processed variables
spotify_charts <- processed_dataframe$spotify_charts
bpm <- processed_dataframe$bpm
danceability <- processed_dataframe$danceability
energy <- processed_dataframe$energy
mode <- processed_dataframe$mode

hist(spotify_charts)
hist(bpm)
hist(danceability)
hist(energy)
hist(mode)

plot(processed_dataframe)
processed_dataframe


bpm_slr <- lm(spotify_charts~bpm)
danceability_slr <- lm(spotify_charts~danceability)
energy_slr <- lm(spotify_charts~energy)

full_model <- lm(spotify_charts~bpm+danceability+energy+factor(mode)
                 +bpm:factor(mode)+danceability:factor(mode)+energy:factor(mode)
                 , qr = TRUE)

```


```{R}
### SLR Model Analysis

#install.packages("lmtest")
library(lmtest)

class(bpm_slr$model$bpm)
class(danceability_slr$model$danceability)
class(energy_slr$model$energy)

### BPM ###
summary(bpm_slr)
anova(bpm_slr)
bpm_slr.residuals <- residuals(bpm_slr)
shapiro.test(bpm_slr.residuals)                           #Reject
bptest(bpm_slr)                                           #Fail to reject
plot(bpm_slr)
hist(bpm_slr$model$bpm)
boxplot(bpm_slr$model$bpm)

### Danceability ###
summary(danceability_slr)
anova(danceability_slr)
danceability_slr.residuals <- residuals(danceability_slr)
shapiro.test(danceability_slr.residuals)                 #Reject
bptest(danceability_slr)                                 #Fail to reject
plot(danceability_slr)
hist(danceability_slr$model$danceability)
boxplot(danceability_slr$model$danceability)

### Energy ###
summary(energy_slr)
anova(energy_slr)
energy_slr.residuals <- residuals(energy_slr)
shapiro.test(energy_slr.residuals)                      #Reject
bptest(energy_slr)                                      #Fail to reject
plot(energy_slr)
hist(energy_slr$model$energy)
boxplot(energy_slr$model$energy )

### Full ###
summary(full_model)
anova(full_model)
full_model.residuals <- residuals(full_model)
shapiro.test(full_model.residuals)
bptest(full_model)
```


```{r}
### Low p-value for the Shapiro-Wilk Test ###
### Apply the Box Cox method ###
# Reminder: Back Transform to interpret model
#install.packages("MASS")
library(MASS)

### BPM ###
y <- bpm_slr$model$spotify_charts
x <- bpm_slr$model$bpm

bc1 <- boxcox(y~x)
(lambda1 <- bc1$x[which.max(bc1$y)])

new_bpm_slr <- lm(((y^lambda1-1)/lambda1) ~ x)

new_bpm_slr.residuals <- residuals(new_bpm_slr)
shapiro.test(new_bpm_slr.residuals)                           #Reject
plot(new_bpm_slr)
hist(new_bpm_slr$model$x)
boxplot(new_bpm_slr$model$x)

# Apply Robust Regression
new_y <- new_bpm_slr$model$`((y^lambda1 - 1)/lambda1)`
new_x <- new_bpm_slr$model$x

new_robust_bpm_slr <- rlm(new_y ~ new_x, psi=psi.bisquare)

new_robust_bpm_slr.residuals <- residuals(new_robust_bpm_slr)
shapiro.test(new_robust_bpm_slr.residuals)                   #Reject
plot(new_robust_bpm_slr)
hist(new_robust_bpm_slr$model$new_x)
boxplot(new_robust_bpm_slr$model$new_x)

### Danceability ###
y <- danceability_slr$model$spotify_charts
x <- danceability_slr$model$danceability

bc2 <- boxcox(y~x)
(lambda2 <- bc2$x[which.max(bc2$y)])

new_danceability_slr <- lm(((y^lambda2-1)/lambda2) ~ x)

new_danceability_slr.residuals <- residuals(new_danceability_slr)
shapiro.test(new_danceability_slr.residuals)                           #Reject
plot(new_danceability_slr)
hist(new_danceability_slr$model$x)
boxplot(new_danceability_slr$model$x)

# Apply Robust Regression
new_y <- new_danceability_slr$model$`((y^lambda2 - 1)/lambda2)`
new_x <- new_danceability_slr$model$x

new_robust_danceability_slr <- rlm(new_y ~ new_x, psi=psi.bisquare)

new_robust_danceability_slr.residuals <- residuals(new_robust_danceability_slr)
shapiro.test(new_robust_danceability_slr.residuals)                   #Reject
plot(new_robust_danceability_slr)
hist(new_robust_danceability_slr$model$new_x)
boxplot(new_robust_danceability_slr$model$new_x)

### Energy ###
y <- energy_slr$model$spotify_charts
x <- energy_slr$model$energy

bc3 <- boxcox(y~x)
(lambda3 <- bc3$x[which.max(bc3$y)])

new_energy_slr <- lm(((y^lambda3-1)/lambda3) ~ x)

new_energy_slr.residuals <- residuals(new_energy_slr)
shapiro.test(new_energy_slr.residuals)                           #Reject
plot(new_energy_slr)
hist(new_energy_slr$model$x)
boxplot(new_energy_slr$model$x)

# Apply Robust Regression
new_y <- new_energy_slr$model$`((y^lambda3 - 1)/lambda3)`
new_x <- new_energy_slr$model$x

new_robust_energy_slr <- rlm(new_y ~ new_x, psi=psi.bisquare)

new_robust_energy_slr.residuals <- residuals(new_robust_danceability_slr)
shapiro.test(new_robust_energy_slr.residuals)                   #Reject
plot(new_robust_energy_slr)
hist(new_robust_energy_slr$model$new_x)
boxplot(new_robust_energy_slr$model$new_x)

# Summary of SLR models
# This gives the impact of each variable
print("Transformed Robust BPM")
summary(new_robust_bpm_slr)
print("Transformed Robust Danceability")
summary(new_robust_danceability_slr)
print("Transformed Robust Energy")
summary(new_robust_energy_slr)
# Factor of 10 improvement on the p-value 
# Still below 0.05
```

```{r}
# MLR analysis
#install.packages("car")
#install.packages("mlr")
#install.packages("boot")
#install.packages("lmridge")
#install.packages("fmsb")
#install.packages("ALSM")
#install.packages("leaps")
#install.packages("caret")
library(car)
library(mlr)

summary(full_model)
Anova(full_model, type = "II")
anova(full_model)
avPlots(full_model)

# Assumption Checking
shapiro.test(full_model.residuals)  #Reject
bptest(full_model)                  #Failed to Reject
plot(full_model)
plot(full_model.residuals)
qqnorm(full_model.residuals)
car::vif(full_model) # Categorical Variable (Mode) presents high levels of 
                   # multicollinearity suggesting that it presents little value 
                   # as a predictor variable 

##################################################
# fmsb as a function overwrite that is not compatible with the 
# previous definition of bptest 
library(fmsb)

# Multicollinearity 
VIF(full_model)     


```

```{r}
#################################################
## Box Cox Transformation
## Handling the non normality
library(MASS)

full__model <- lm(spotify_charts~bpm+danceability+energy) #Double underscore

bc_full <- boxCox(full__model, lambda = seq(-2,2,0.1))
lambda_full <- bc_full$x[which.max(bc_full$y)]

boxcox_spotify_charts <- ((spotify_charts^lambda_full-1)/lambda_full)


#class(boxcox_spotify_charts)
#length(boxcox_spotify_charts)

full_boxcox_model <- lm(boxcox_spotify_charts~bpm+danceability+energy+factor(mode)
                 +bpm:factor(mode)+danceability:factor(mode)+energy:factor(mode))


summary(full_boxcox_model)
Anova(full_boxcox_model, type = "II")
anova(full_boxcox_model)
full_boxcox_model.residuals <- residuals(full_boxcox_model)
shapiro.test(full_boxcox_model.residuals)                  #Reject
plot(full_boxcox_model)                                    #Again large improvement but still not passing
plot(full_boxcox_model.residuals) 

#################################################
# Check for influential points
# Handling the non normality continued

DFFITS = dffits(full_model)
Cook = cooks.distance(full_model)
DFBETAS = dfbetas(full_model)
influential_points_dfbetas <- which(abs(DFBETAS) > 2, arr.ind = TRUE)
influential_points_cooks <- which(Cook > 4 / nobs(full_model), arr.ind = TRUE)
avg_dffits <- sqrt(mean(DFFITS^2))
influential_points_dffits <- which(abs(DFFITS) > 2 * avg_dffits, arr.ind = TRUE)

#influential_points_cooks
#influential_points_dfbetas
#influential_points_dffits

plot(full_model, pch = 18, col = "red", which = c(4))
dfbetasPlots(full_model)
influencePlot(full_model)


# Influential Point Output
# 1   3   4   6   9  13  14  15  16  18  23  25  30  31  33  42  43  44  45  49  55  56  68 361 452 470 

robust_full_model <- rlm(spotify_charts~bpm+danceability+energy+factor(mode)
                         +bpm:factor(mode)+danceability:factor(mode)
                         +energy:factor(mode),
                         data=processed_dataframe, psi=psi.bisquare)
robust_full_model.residuals <- residuals(robust_full_model)
summary(robust_full_model)
shapiro.test(robust_full_model.residuals)


y <- full_boxcox_model$model$boxcox_spotify_charts
x1 <- full_boxcox_model$model$bpm
x2 <- full_boxcox_model$model$danceability
x3 <- full_boxcox_model$model$energy
x4 <- full_boxcox_model$model$`factor(mode)`


robust_full_boxcox_model <- rlm(y ~ x1 + x2 + x3 + x4 + x1:x4 + x2:x4 + x3:x4,
                                psi=psi.bisquare)
robust_full_boxcox_model.residuals <- residuals(robust_full_boxcox_model)
summary(robust_full_boxcox_model)
shapiro.test(robust_full_boxcox_model.residuals)
plot(robust_full_boxcox_model)

summary(robust_full_boxcox_model)
anova(robust_full_boxcox_model)
##NOTICE:
#Another Major improvement in the p-value 
#Still below 0.05 @ 0.0001285
```


```{r}
library(lmridge)
## Addressing Multicollinearity

avPlots(robust_full_boxcox_model)
avPlots(full__model)

car::vif(robust_full_boxcox_model)
# Even after addressing normality the categorical variable is showing 
# High levels of multicollinearity

#Ridge Method on Transformed Full Model 
robust_full_boxcox_model.ridge <- lm.ridge(robust_full_boxcox_model, data = processed_dataframe, lambda = seq(0,100,0.02))
#robust_full_boxcox_model.ridge <- lm.ridge(robust_full_boxcox_model, data = processed_dataframe, lambda = seq(0,1000,0.02))
plot(robust_full_boxcox_model.ridge)
MASS::select(robust_full_boxcox_model.ridge)


car::vif(full__model)
#Ridge Method model without factor(mode)
full__model.ridge <- lm.ridge(full__model, data = processed_dataframe, lambda = seq(0,10,1))
#full__model.ridge <- lm.ridge(full__model, data = processed_dataframe, lambda = seq(0,10000,1))
plot(full__model.ridge)
MASS::select(full__model.ridge)

## Each run of the ridge regression on both the model (with & without factor(mode))
## The GCV that is related to lambda is the max value of the sequence 
## Since Lambda manages the shrinkage of the distribution
## You would want to balance the size of lambda with the amount of variance 
##  that is described by the predictor variable, 
## (i.e. the max might not be as desirable as the sequence grows to infinity)
## What this tells me is that there is some amount of multicollinearity that 
##  cannot be removed without trivializing the model 


```


```{r}

################################################
# Validation
library(caret)
library(leaps)
# Compare the Robust Full BoxCox and the Full Model to compare results
set.seed(42)
train.control = trainControl(method="cv", number = 10)

step.model1 = train(spotify_charts~bpm+danceability+energy+factor(mode)
                 +bpm:factor(mode)+danceability:factor(mode)+energy:factor(mode),
	data = processed_dataframe, method = "leapBackward", tuneGrid=data.frame(nvmax=7), 
	trControl = train.control)
step.model1$results

step.model2 = train(y ~ x1 + x2 + x3 + x4 + x1:x4 + x2:x4 + x3:x4,
                    data = data.frame(robust_full_boxcox_model$model),	
                    method = "leapBackward", tuneGrid=data.frame(nvmax=7),
	                  trControl = train.control)
step.model2$results


```